### 2023-09-29 documentation/thesis/thesis_declaration.tex

I hereby declare that I have prepared this thesis independently and
without outside assistance. Text passages, which are based literally or
in the sense on publications or lectures of other authors, are marked as
such. The work has not yet been submitted to any other examination
authority and has not yet been published.

Aachen, \datum  \_\_\_\_\_\_\_\_\_\_\_\_,
\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_