### 2023-10-03 documentation/thesis/thesis_abstract.tex
.

Especially in the construction of low-field MRI devices based on
permanent magnets, a large number of magnets are used. In order to
realize a homogeneous B0 field with these magnets, which is necessary
for many setups, the magnetic properties of these magnets should be as
similar as possible.

Due to the complex manufacturing process, especially of neodymium
magnets, the different properties, especially the direction of
magnetization, can deviate from each other, which affects the
homogeneity of the field.

To adjust the field afterwards, a passive shimming process is typically
performed, which is complex and time-consuming and requires manual
corrections to the magnets used.

To avoid this process, magnets can be systematically measured in
advance. However, in this methodology, the recording, data storage and
subsequent evaluation of the data play an important role.

The various existing solutions implement individual aspects, but do not
provide a data pipeline from aqusation to analysis.

For this use case, the MagneticReadoutProcessing library was created,
which implements different aspects of data acquisition, data storage,
analysis, and each intermediate step can be customized by the user
without having to create everything from scratch, favoring an exchange
between different user groups.