### 2024-01-07 documentation/thesis/thesis_abstract.tex
more thesis - first external review

A large number of magnets are used in the construction of low-field MRI
equipment on the basis of permanent magnets. The magnetic properties of
these magnets must be similar to a certain degree in order to achieve a
homogeneous B0 field, which is necessary for many setups. Due to the
complex manufacturing process of neodymium magnets, the different
properties, i.e.~the direction of magnetisation, can deviate from each
other. This affects the homogeneity of the field. A passive shimming
process is typically used to adjust the field afterwards. This is
complex and time-consuming and requires manual corrections to the
magnets used. To avoid this process, magnets can be systematically
measured in advance. Data acquisition, storage and subsequent analysis
play an important role in this methodology. Several existing open source
solutions implement individual parts, but do not provide a complete data
processing pipeline from acquisition to analysis, and their data storage
formats are not compatible with each other. For this use case, the
MagneticReadoutProcessing library has been created. It implements all
important aspects of acquisition, storage and analysis, and each
intermediate step can be customised by the user without having to create
everything from scratch, thus encouraging exchange between different
user groups. Complete documentation, tutorials and tests enable users to
use and adapt the framework as quickly as possible. The framework was
used to characterise different magnets, which requires integrating
magnetic field sensors.