### 2024-02-27 documentation/colloquium/presentation/index.qmd
.

---
title: "Development of a permanent magnet characterisation framework for use in low-field MRI systems"
subtitle: "FH Aachen – University of Applied Sciences"
author: Marcel Ochsendorf
institute: Faculty of Electrical Enginnering and Information Technology - Information Systems Engineering
date: today
date-format: "dddd, [the] D[<sup style='font-size:65%;font-style:italic;'>th</sup>] [of] MMMM, YYYY"
embed-resources: false
format: coeos-revealjs
---

```{r}
#| include: false
sysfonts::font_add_google("Alegreya Sans", "Alegreya Sans", regular.wt = 300)
showtext::showtext_auto()
source("_extensions/mcanouil/coeos/setup-ggplot2-coeos.R")
```

## Introduction

* test

## Aim of this Project

---

##

<h1 class="r-fit-text" style="text-align: center;">MagneticReadoutProcessing</h1>

<p class="fragment">Provides and user-oriented **open** magnetic field data hardware and software framework:</p>
<br />
<p class="fragment">&#x2022; Harware support for several magnetic field sensors</p>
<br />
<p class="fragment">&#x2022; Easy to integrate into existing projects</p>
<br />
<p class="fragment">&#x2022; Easy to extend</p>

<aside class="notes">
    *
    Folgend ein paar beispiele
</aside>

---

### Hardware Framework

::: {.container}
:::: {.col}
<h4 style="text-align: center;">3D Rotational Sensor</h4>
::::
:::

::: {.container}
:::: {.col}
<img src="images/3D_sensor.png" alt="image" height="320px">
::::

:::: {.col}
<img src="images/3D_sensor_measurement.png" alt="image" height="320px" >
::::
:::

---

### Hardware Framework

::: {.container}
:::: {.col}
<h4 style="text-align: center;">1D Static Sensor</h4>
::::
:::

::: {.container}
:::: {.col}
<img src="images/1D_sensor.png" alt="image" height="320px">
::::

:::: {.col}
<img src="images/1D_sensor_measurement.png" alt="image" height="320px" >
::::
:::

---

### Hardware Framework

Cheap and easy to construct using consumer grade magnetic field sensors:
<br />
<p class="fragment">&#x2022; One software for most common digital magnetic field sensors</p>
<br />
<p class="fragment">&#x2022; Synchronised multi sensor setups</p>
<br />
<p class="fragment">&#x2022; Universal and interachangable magnet mount</p>

<aside class="notes">
    In der software sind aktuelll 8 typische sensoren hinterlegt.
    Magnet mount für viele Magnetformen anpassbar, ohne dass die sensor hardware geaendert werden muss
</aside>

---

### Software Framework

Simply tasks of:

*
* data aqusition
* data storage
* data analysis tasks

---

### User Interaction Points

![](images/MRP_high_level_overview.png)

---

### User Interaction Points

Allows rapid development of software analysation functions, witout implementation of typical data processing pipelines needed by providing:

<br />
<p class="fragment">&#x2022; Expandable open datatypes for magnetic field data</p>
<br />
<p class="fragment">&#x2022; Prebuild pipelines ready for customisation</p>
<br />
<p class="fragment">&#x2022; Full online documentation, tested library and user tutorials</p>

---

## Use Cases

Two pratical application of the framework will be shown using two examples:

<br />
<p class="fragment">&#x2022; Characterization of a *MM5603NJ* and determination of background noise and temperature linearity</p>
<br />
<p class="fragment">&#x2022; Usage of the chrakterised sensor for sorting permanent magnets</p>

<aside class="notes">
    Es wurden wärend der Arbeit weitere Sensoren chrakterisiert, es wird jedoch nur dieser sensor als beispiel genommen.
    Mittels der chraktisierung im vorfeld ist eine genauerer chrakterisierung der Magnete möglich.

</aside>
---

### Sensor Chraterisation

![](images/Sensor_evaluation_setup_for_linearity_measurements.png)

<aside class="notes">
    Grafik zeigt den vereinfachten Aufbau welcher für die Chrakterisierung verwendet wurde.
    Platte zeigen, verschiedene Sensoren wurden unter den geleichen bedingungen vermessen.
</aside>

---

### *MM5603NJ* chraterisation results

---

### Manget Sorting

<aside class="notes">
    Die Sortierung wird anhand einer Demo des Frameworks durchgeführt.
    Dabei werden die sensor charakterisiungsdaten verwendet
    Hierbei wird eine preconfiguierte Pipeline verwendet.
    Das ergebnis wird am ende eine reduktion der vermessenen Magnete sein.
    Somit kann dann ein optimierter hallbach ring generiert werden um so die homogentität zu verbessern
</aside>

---

#### DEMO: MRPCli

* Generate automated measurement using hardware sensors
* Save results to filesystem for later analysation

![](images/MRPCli.gif)

---

#### DEMO: MRPUdpp

* Execution of pre-configured analysation and pipeline
* User implements filter algorithm only

![](images/MRPUdpp.gif)

---

## Conclusion

---

# 
<h1 class="r-fit-text" style="text-align: center;">Questions ?</h1>

<table>
<tr><td><p>[<img src="images/github-mark-white.png" alt="image" width="150" height="auto">](https://github.com/LFB-MRI/MagneticReadoutProcessing) </p></td><td><p>[<img src="images/researchgate.svg" alt="image" width="150" height="auto">](https://www.researchgate.net/publication/374388764_Development_of_a_permanent_magnet_characterisation_framework_for_use_in_low-field_MRI_systems) </p><td>[<img src="images/avatar.png" alt="image" width="150" height="auto">](https://marcelochsendorf.com)</td><td><p>[<img src="images/pypi.svg" alt="image" width="150" height="auto">](https://pypi.org/project/MagneticReadoutProcessing/) </p></td><td><p>[<img src="images/rtd.png" alt="image" width="150" height="auto">](https://magneticreadoutprocessing.readthedocs.io/en/latest/) </p></td></tr>
</table>